---
title: "CS 1675 Fall 2020: Coding live stream examples"
subtitle: "Debugging examples"
author: "Dr. Joseph P. Yurko"
date: "9/28/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This markdown includes errors throughout several code chunks to practice debugging. The code chunks are all set to `eval=TRUE` for now, it is not possible to knit the document. If you want to knit and render the HTML file to just look at the code without any output, place `eval=FALSE` in the "document level" chunk options controlled by `knitr::opts_chunk()` in the `setup` code chunk.  

```{r, load_packages}
library(tidyverse)
```

Define a function which has a step that has a matrix multiplication.  

```{r, define_my_function}
my_function <- function(beta, my_info)
{
  X <- my_info$the_matrix_x
  
  beta * X
}
```

Define a matrix `X` that is 5 rows and 3 columns.  

```{r, make_the_x_matrix}
set.seed(7123132)
x1 <- rnorm(n = 5)
x2 <- rnorm(n = 5)

Xmat <- cbind( `(Intercept)` = rep(1, length(x1)), 
               x1 = x1, 
               x2 = x2)

Xmat
```

Pass in the matrix `Xmat` into the `my_function()` function and multiply by a scalar number, 3. First, set the list of required information.  

```{r, set_the_info_to_use}
info_use <- list(
  the_matrix_x = Xmat
)
```

Now, call the function.  

```{r}
my_function(beta = 3, my_info = info_use)
```

What if I used a vector for the `beta` input argument?  

```{r, set_a_beta_vector}
beta_try <- c(1.25, 2.5)

length(beta_try)

dim(beta_try)

class(beta_try)
```

Try the multiplication with the `beta_try` vector.  

```{r}
my_function(beta_try, info_use)
```

Element-wise multiplication, multiplies each element of two objects together.  

```{r}
c(1, 2, 3, 4) * c(4, 3, 2, 1)
```

We can manually perform the above operation as:  

```{r}
y1 <- 1:4
y2 <- rev(1:4)
```

```{r}
c(y1[1] * y2[1], y1[2] * y2[2], y1[3] * y2[3], y1[4] * y2[4])
```

The inner product requires a dimensionality because it is between matrices. Where a row or column vector is a 1D matrix.  

Let's now convert the regular vector into a matrix object.  

```{r}
as.matrix(1:4)

matrix(1:4)
```

The `matrix()` function has a few more options for control. For example to create a $2 \times 2$ matrix we can set `nrow=2`.  

```{r}
matrix(1:4, nrow = 2)
```

And we can choose to "fill" the matrix by rows or columns, first.  

```{r}
matrix(1:4, nrow = 2, byrow = TRUE)
```

Let's define a regular vector, `beta_vec`, that we then convert to a matrix type column vector, `beta_col`.  

```{r}
beta_vec <- c(1.25, -1, 2)

beta_col <- matrix(beta_vec)

beta_col
```

What will happen if I run `my_function()` with `beta_col` as the `beta` argument?  

```{r}
my_function(beta_col, info_use)
```

Check the data types and dimensionalities.  

```{r}
class(info_use$the_matrix_x)

dim(info_use$the_matrix_x)

class(beta_col)

dim(beta_col)
```

Let's try and debug our code. To save space, we will define a new function that does the same thing as `my_function()`.  

Debug using the `browser()` function. Or "hack" a debug with forcing a `return()`.  

```{r}
same_as_my_function <- function(beta, my_info)
{
  X <- my_info$the_matrix_x
  
  # browser()
  
  # return(class(X %*% beta))
  
  # return(dim(beta))
  
  beta * X
}
```

Run `same_as_my_function()` with the input arguments that created the error.  

```{r}
same_as_my_function(beta_col, info_use)
```

Code the correct function and call it, `my_mean_trend()`.  

```{r}
my_mean_trend <- function(beta, my_info)
{
  X <- my_info$the_matrix_x
  
  X %*% beta
}
```

```{r}
my_mean_trend(beta_col, info_use)
```

We can convert the result of `my_mean_trend()` from a matrix to a "regular" vector using `as.numeric()`.  

```{r}
class( my_mean_trend(beta_col, info_use) )

as.numeric( my_mean_trend(beta_col, info_use) )
```

What would happen if I pass in a column vector with the wrong number of elements?  

```{r}
my_mean_trend(as.matrix(1:7), info_use)
```

What happens if we pass in a "regular" numeric vector?  

```{r}
my_mean_trend(beta_vec, info_use)
```

Check that this is the same result as before.  

```{r}
my_mean_trend(beta_col, info_use) - my_mean_trend(beta_vec, info_use)
```

```{r}
all.equal(my_mean_trend(beta_col, info_use), my_mean_trend(beta_vec, info_use))
```

```{r}
info_use$the_matrix_x
```

```{r}
t(beta_col) %>% as.numeric()
```

```{r}
beta_col %>% as.numeric()
```

```{r}
all.equal(as.numeric(beta_col), beta_vec)
```

```{r}
all.equal(as.numeric(t(beta_col)), beta_vec)
```

What about a matrix?  

```{r}
as.numeric(matrix(1:4, nrow = 2))
```

```{r}
matrix(1:4, nrow=2)
```

Compare that to when the matrix is ordered by rows.  

```{r}
as.numeric(matrix(1:4, nrow = 2, byrow = TRUE))
```

```{r}
matrix(1:4, nrow=2, byrow=TRUE)
```

